\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-23 trunk/europa.tex
Corrected missed references

\eu is implemented as a library that encapsulates all the elements
described above and makes them easily accessible through modeling and
Application Programming Interface (API) layers.  A user will typically
use the modeling facilities to inject a problem domain description,
and the API to extract and manipulate plans, and to obtain information
about plan flaws and constraint violations at any stage of the
planning process.

@@

Fig. \ref{fig:europa-architecture} shows the main architectural
components in \eu and their relationships.

The \textbf{Constraint Reasoning Engine} (CRE) manages domain types,
variables and constraints that define relationships among them. It
also provides an efficient arc consistency mechanism
\cite{mackworth77}. The CRE is designed so that specialized reasoning
algorithms for new constraints can be easily and efficiently plugged
in.

The \textbf{Plan Database} (PDB) manages object and token types. It
also maintains the state of partial plans in terms of object, variable
and token instances. This constitutes the data foundation on which end
user applications and automated planners are built.

The \textbf{Rules Engine} (RE) modularly adds the semantics of domain
rules that describe dependencies between tokens.

The \textbf{Solver Module} provides the framework for flaw detection
and resolution that is needed to implement search algorithms for
planning.

The \textbf {Language Modules}: Describing problem domains and problem
instances in terms of objects, variables and domain rules in a
programmatic way is cumbersome to the extent of being impractical,
this motivated the creation of modeling languages that support those
definitions in a much more natural and compact way. \texttt{NDDL} (New
Domain Description Language [pronounced ``noodle'']) \cite{NDDL} is
the main modeling language for \eu and will be described in detail in
the Modeling section below. ANML is a language under development that
provides a more intuitive syntax \cite{smith08}.

The \textbf{API Layer} provides access to services from all the
different modules in a programmatic way so that \eu can be easily
embedded into specialized applications.

@@

\item \textit{Temporal Network}: Extensions to the CRE module to
  reason about temporal constraints

\item \textit{Resources}: Extensions to the CRE, PDB and Solver
  modules to provide out-of-the-box reasoning and search for metric
  resources.

\item \textit{Chronological Backtracking Solver}: A general purpose
  built-in solver to provide the ability to plan out-of-the-box,
  problems that require challenging search may require specialized
  solvers to be built as explained in the Search section below.
\end{enumerate}

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eue,
\texttt{NDDL} allows the user to specify the different components of a
problem domain in a concise way. The main characteristics that make
\texttt{NDDL} a powerful tool for describing problem domains are:

\item Object Oriented: \texttt{NDDL} supports classes and inheritance
  in similar fashion to popular Object-Oriented languages like Java
  and C++. It also supports polymorphism for some of the planning
  components. Using object classes and instances is a time-tested
  approach to naturally describe problem domains \cite{rumbaugh}.

\item It offers declarative constructs to define constraints and
  causality dependencies between actions and effects.

\item It offers procedural constructs to populate \eu's plan database
  with specific problem instances expressed in terms of objects,
  variables, constraints, facts and goals.

To understand how the most important elements of \texttt{NDDL} work,
we show how the Shopping Agent problem could be modeled:

@@

A problem instance consists of:

\item \textbf{Variables} and \textbf{Objects}: In this case we defined
  temporal variables that are used to specify the planning horizon
  (see \texttt{horizonStart} and \texttt{horizonEnd}), and objects are
  used to specify the locations, products, product-location
  associations and the Shopping Agent that we want to plan for.

\item \textbf{Facts}: These are used to specify the current state of
  the world and don't need any support. Facts are represented as
  tokens in \eu's Plan Database. In this case we use facts to specify
  that the Shipping Agent is initially at home.

\item \textbf{Goals}: Goals are represented as tokens in \eu's Plan
  Database. Differently from Facts, Goals trigger inference in \eu
  which determines whether flaws (unresolved subgoals, overloaded
  resources, etc.) need to be resolved, if that is the case, search
  modules can be invoked which will look for a plan to resolve those
  flaws.

\item \textbf{Constraints}: restrictions on and between actions and
  goals are specified through constraints, in this case we use
  temporal constraints to specify when the facts hold and when the
  goals should be accomplished. As always, all the constraints in
  \eu's library (built-in, or user defined), can be used to specify
  these restrictions.
\end{enumerate}

@@

In \eue, all the Variables and Constraints from a problem domain,
domain descriptions are kept inside the Constraint Reasoning Engine
(CRE).  The CRE provides a framework where any number of constraint
types can collaborate to perform inference and prune variable domains
using Bounds Propagation and Domain Reduction techniques. \eu uses a
Propagator pattern, in which a Propagator object keeps track of a set
of constraint instances of the same type and is responsible for
ensuring that whenever any variable changes or constraints are
introduced or dropped, all related variables are updated through
inference.

In a planning problem, variable or constraint changes can come
directly from an end user (by introducing new activities or goals and
relating them to the rest of the plan through constraints, by changing
the start time of an activity, etc), or can come from an automated
planner (by expanding subgoals, introducing precedence constraints to
resolve mutual exclusion conflicts, etc). When these changes happen,
each Propagator is only able to ensure consistency for the constraints
that are associated with it.  To maintain global consistency, the CRE
implements a variant of the \texttt{AC-3} algorithm \cite{mackworth77}
where it asks Propagators to perform inference until the Constraint
Network returns to an stable state.

%\kcomment{add europa's AC-3 algorithm here}

@@

Temporal and resource constraints are specially important to model
real-life planning domains, so it is worth going into some more detail
about how \eu's built-in support works for those constraint types.

@@

>>>>

\eu is implemented as a library that encapsulates all elements
described above making them accessible via modeling and Application
Programming Interface (API) layers.  A user will typically use the
modeling facilities to inject a problem domain description and the
API to extract and manipulate plans and to obtain information about
plan flaws and constraint violations at any stage of the planning
process.

@@

Fig. \ref{fig:europa-architecture} shows the main architectural
components in \eu and their relationships. The \textbf{Constraint
  Reasoning Engine} (\texttt{CRE}) manages domain types, variables and
constraints that define relationships among them. It also provides an
efficient arc consistency mechanism \cite{mackworth77}. The
\texttt{CRE} is designed so that specialized reasoning algorithms for
new constraints can be easily and efficiently plugged in. The
\textbf{Plan Database} (\texttt{PDB}) manages object and token
types. It also maintains the state of partial plans in terms of
object, variable and token instances. This constitutes the data
foundation on which end user applications and automated planners
within \eu are built. The \textbf{Rules Engine} (\texttt{RE})
modularly adds the semantics of domain rules that describe
dependencies between tokens. The \textbf{Solver Module} provides the
framework for flaw detection and resolution that is needed to
implement search algorithms for planning. The \textbf{API Layer}
provides access to services from all the different modules in a
programmatic way so that \eu can be easily embedded into specialized
applications.

Describing problem domains and problem instances in terms of objects,
variables and domain rules in a programmatic way is cumbersome to the
extent of being impractical; this motivated the creation of modeling
languages that support those definitions in a much more natural and
compact way. \texttt{NDDL} (New Domain Description Language
[pronounced ``noodle'']) \cite{NDDL} is the main modeling language for
\eu and will be described in detail in the Modeling section
below. Another, \texttt{ANML} is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

\item \textit{Temporal Network}: Extensions to the \texttt{CRE} module to
  reason about temporal constraints

\item \textit{Resources}: Extensions to the \texttt{CRE}, \texttt{PDB}
  and Solver modules to provide reasoning and search for metric
  resources.

\item \textit{Chronological Backtracking Solver}: A general purpose
  built-in solver to provide the ability to plan out-of-the-box.
  Problems that require challenging search may require specialized
  solvers to be built as explained in Section \ref{sec:europa:search}.

@@

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eue,
\texttt{NDDL} allows the user to specify the different components of a
problem domain in a precise and concise manner. Among the main
features that make \texttt{NDDL} a powerful tool for describing
problem domains are that it is object oriented. \texttt{NDDL} supports
classes and inheritance in similar fashion to popular Object-Oriented
languages like Java and C++. It also supports polymorphism for some of
the planning components. Using object classes and instances is a
time-tested approach to naturally describe problem domains
\cite{rumbaugh}. It offers declarative constructs to define
constraints and causality dependencies between actions and effects and
finally it offers procedural constructs to populate \eus plan
database with specific problem instances expressed in terms of
objects, variables, constraints, facts and goals. To articulate
important elements of \texttt{NDDL}, we show how the Shopping Agent
problem could be modeled:

@@

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts are used to specify the current state of the
world and do not need any support and are represented as tokens in
\eus plan database. In this case we use facts to specify that the
Shipping Agent is initially at home. Goals are also represented as
tokens in \eus plan database. Different from facts, goals trigger
inference in \eu which determines whether flaws (unresolved subgoals,
overloaded resources, etc.) need to be resolved; if so, search modules
can be invoked to resolve those flaws as part of the deliberation
process. Finally, constraints are restrictions on and between actions
and goals. In the example above, we use temporal constraints to
specify when facts hold and when the goals should be
accomplished. % As
% always, all constraints in \eus library could be used.

@@

In \eue, all variables and constraints from a problem domain
description are stored inside the \texttt{CRE}.  The \texttt{CRE}
provides a framework where any number of constraint types can
collaborate to perform inference and prune variable domains using
Bounds Propagation and Domain Reduction techniques \kcomment{citation
  needed}. \eu uses a Propagator pattern which keeps track of a set of
constraint instances of the same type and is responsible for ensuring
that whenever any variable is modified or constraints are introduced
or dropped, all related variables are updated through inference.

In a planning problem, variable or constraint changes can come
directly from an end user (by introducing new activities or goals and
relating them to the rest of the plan through constraints, by changing
the start time of an activity, etc), or can come from an automated
planner (by expanding subgoals, introducing precedence constraints to
resolve mutual exclusion conflicts, etc). When these changes happen,
each Propagator is only able to ensure consistency for the constraints
that are associated with it.  To maintain global consistency, the
\texttt{CRE} implements a variant of the \texttt{AC-3} algorithm
\cite{mackworth77} where it asks Propagators to perform inference
until the Constraint Network returns to an stable state.

@@

Temporal and resource constraints are specially important to model
real-life planning domains, so it is worth going into some more detail
about how \eus built-in support works for those constraint types.

@@

reflected by the use of the \texttt{Allen Algebra} \cite{allen84}
temporal relations shown in Fig. \ref{fig:allen-algebra}. Interval
arithmetic in the EUROPA$_2$ domain model using \texttt{Allen Algebra}
allows encoding of temporal constraint relationships between tokens on
concurrent timelines. See \cite{jonsson00} for additional details
including propagation algorithms and interval logic.
